---
title: "Slides"
format: revealjs
bibliography: references.bib
---

## Set-up libs


```{r}
#| label: necessary-packages

library(tidyverse)
library(sf)
library(gt)
library(tmap)
library(srvyr)
library(arrow)
```




# Introduction to the European Social Survey (ESS)

## Overview 

- Conducted across Europe every 2 years since 2001
- Cross-sectional sample (new samples each round)
- Topics include living conditions, social structure, public opinion and attitudes
- Sampling designs vary by country but have oversight for rigor

## Countries for round 11 (2023-2024)

```{r}
#| label: load-ess
#| cache: true
#| echo: false

url <- glue::glue(
  "https://api.ess.sikt.no/v1/data/dataFile/10.21338/ess11e04_2?userId={Sys.getenv('ESS_USERID')}"
)
ess_in <- read_parquet(url)
```

```{r}
#| label: round-11-map
#| echo: false
country_coding <-
  tibble::tribble(
    ~Code , ~Name                ,
    'AL'  , 'Albania'            ,
    'AT'  , 'Austria'            ,
    'BE'  , 'Belgium'            ,
    'BG'  , 'Bulgaria'           ,
    'CH'  , 'Switzerland'        ,
    'CY'  , 'Cyprus'             ,
    'CZ'  , 'Czechia'            ,
    'DE'  , 'Germany'            ,
    'DK'  , 'Denmark'            ,
    'EE'  , 'Estonia'            ,
    'ES'  , 'Spain'              ,
    'FI'  , 'Finland'            ,
    'FR'  , 'France'             ,
    'GB'  , 'United Kingdom'     ,
    'GE'  , 'Georgia'            ,
    'GR'  , 'Greece'             ,
    'HR'  , 'Croatia'            ,
    'HU'  , 'Hungary'            ,
    'IE'  , 'Ireland'            ,
    'IS'  , 'Iceland'            ,
    'IL'  , 'Israel'             ,
    'IT'  , 'Italy'              ,
    'LT'  , 'Lithuania'          ,
    'LU'  , 'Luxembourg'         ,
    'LV'  , 'Latvia'             ,
    'ME'  , 'Montenegro'         ,
    'MK'  , 'North Macedonia'    ,
    'NL'  , 'Netherlands'        ,
    'NO'  , 'Norway'             ,
    'PL'  , 'Poland'             ,
    'PT'  , 'Portugal'           ,
    'RO'  , 'Romania'            ,
    'RS'  , 'Serbia'             ,
    'RU'  , 'Russian Federation' ,
    'SE'  , 'Sweden'             ,
    'SI'  , 'Slovenia'           ,
    'SK'  , 'Slovakia'           ,
    'TR'  , 'Turkey'             ,
    'UA'  , 'Ukraine'            ,
    'XK'  , 'Kosovo'
  )

ess_11_countries <-
  ess_in |>
  distinct(cntry) |>
  mutate(
    Country = recode_values(
      cntry,
      from = country_coding$Code,
      to = country_coding$Name
    )
  )

country_shape_col <-
  World |>
  # filter(continent %in% c("Europe", "Asia")) |>
  full_join(ess_11_countries, by = c("name" = "Country")) |>
  st_transform(3035) |>
  st_crop(
    st_bbox(c(xmin = -10, ymin = 32, xmax = 36, ymax = 70), crs = 4326) |>
      st_transform(3035)
  ) |>
  mutate(ESS11 = !is.na(cntry))

country_shape_col |>
  tm_shape() +
  tm_polygons(fill = "lightgrey", col = "darkgrey") +
  tm_shape(country_shape_col |> filter(ESS11)) +
  tm_polygons(
    fill = "MAP_COLORS",
    fill.scale = tm_scale_categorical(
      values = c("#0b3954", "#087e8b", "#bfd7ea", "#ff8484", "#8d6b94")
    )
  ) +
  tm_text("cntry") +
  tm_graticules()
```

## Accessing ESS data

1. Create an account at the [ESS Data Portal](https://ess.sikt.no/en/register?redirectUrl=/en/) and then get your [user ID](https://ess.sikt.no/en/api)

2. Add your user ID to your R environ file by opening the file and editing as follows:

```r
usethis::edit_r_environ()
```

Edit as follows (example):

```r
ESS_USERID="12345678-1234-1234-1234-12345678900"
```

3. Restart your R session



## Load the ESS data into R

```{r}
#| label: rep-load-ess
#| ref.label: "load-ess"
#| echo: true
#| eval: false
```

## Why weight? Why use the design?

- Differential selection probabilities and nonresponse are accounted for in analytic weights
- Weights are post-stratified by gender, age, education, and geographic region
- Not using weights leads to bias
- Not using design likely leads to incorrect precision, wrong degrees of freedom and not accounting for clustering

## Estimates changing with weights

```{r}
#| label: why-weight
#| echo: false
ess_der_wt_examp <-
  ess_in |>
  select(
    psu,
    cntry,
    stratum,
    anweight,
    pspwght,
    pdwrk:dngna,
    mainact
  ) |>
  mutate(
    across(
      pdwrk:dngoth,
      ~ if_else(dngdk == 1 | dngref == 1 | dngna == 1, NA, .x),
      .names = "{.col}_rec"
    ),
    Retired = case_when(
      # main activity
      mainact < 66 ~ mainact == 6,
      mainact == 66 ~ rtrd == 1,
      TRUE ~ NA
    ),
    Country = recode_values(
      cntry,
      from = country_coding$Code,
      to = country_coding$Name
    ),
    wt1 = 1
  )

ess_des <- ess_der_wt_examp |>
  as_survey_design(ids = psu, strata = stratum, weights = anweight, nest = TRUE)
ess_des_1 <- ess_der_wt_examp |>
  as_survey_design(ids = psu, strata = stratum, weights = wt1, nest = TRUE)
ess_des_wt <- ess_der_wt_examp |>
  as_survey_design(weights = anweight)
ess_des_srs <- ess_der_wt_examp |>
  as_survey_design(weights = wt1)

calc_ests <- function(des) {
  des |>
    summarize(
      .by = Country,
      p = survey_mean(
        Retired,
        proportion = TRUE,
        prop_method = "logit",
        vartype = "ci",
        na.rm = TRUE
      )
    ) |>
    rename(p_est = p)
}

allests <-
  list(a = ess_des, c = ess_des_wt, b = ess_des_1, d = ess_des_srs) |>
  map(calc_ests) |>
  bind_rows(.id = "method")
```

```{r}
#| label: weight-compare-plot
#| echo: false
allests |>
  mutate(
    Design = recode_values(
      method,
      "a" ~ "Weights, strata, clusters",
      "b" ~ "Strata and clusters only",
      "c" ~ "Weights only",
      "d" ~ "No design"
    ) |>
      factor(
        levels = c(
          "Weights, strata, clusters",
          "Weights only",
          "Strata and clusters only",
          "No design"
        )
      )
  ) |>
  ggplot(aes(color = Design)) +
  geom_point(aes(x = Design, y = p_est)) +
  geom_errorbar(aes(x = Design, ymin = p_low, ymax = p_upp)) +
  theme_light() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.spacing = unit(0, "in")
  ) +
  facet_wrap(~Country) +
  labs(
    title = "Retirement is main activity in last 7 days, ESS 11",
    y = "Estimated proportion with 95% CI"
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, max(allests$p_upp))
  ) +
  scale_color_manual(values = c("#0b3954", "#087e8b", "#ff8484", "#8d6b94"))
```

## ESS Variable Viewer

- Explore ESS variables at the [Variable Viewer](https://ess.sikt.no/en/datafile/242aaa39-3bbb-40f5-98bf-bfb1ce53d8ef?tab=1)

# Closing

## References

---
nocite: |
  @*
---

