---
title: "Slides"
format: revealjs
bibliography: references.bib
---

## Set-up libs


```{r}
#| label: neccessary-packages

library(tidyverse)
library(sf)
library(gt)
library(tmap)
library(srvyr)
library(arrow)
```




# Introduction to the European Social Survey (ESS)

## Overview 

- Conducted across Europe every 2 years since 2001
- Cross-sectional sample (new samples each round)
- Topics include living conditions, social structure, public opinion and attitudes
- Sampling designs vary by country but have oversight for rigor

## Countries for round 11 (2023-2024)

```{r}
#| label: load-ess
#| cache: true
#| echo: false

url <- glue::glue(
  "https://api.ess.sikt.no/v1/data/dataFile/10.21338/ess11e04_2?userId={Sys.getenv('ESS_USERID')}"
)
ess_in <- read_parquet(url)
```

```{r}
#| label: round-11-map
#| echo: false
country_coding <-
  tibble::tribble(
    ~Code , ~Name                ,
    'AL'  , 'Albania'            ,
    'AT'  , 'Austria'            ,
    'BE'  , 'Belgium'            ,
    'BG'  , 'Bulgaria'           ,
    'CH'  , 'Switzerland'        ,
    'CY'  , 'Cyprus'             ,
    'CZ'  , 'Czechia'            ,
    'DE'  , 'Germany'            ,
    'DK'  , 'Denmark'            ,
    'EE'  , 'Estonia'            ,
    'ES'  , 'Spain'              ,
    'FI'  , 'Finland'            ,
    'FR'  , 'France'             ,
    'GB'  , 'United Kingdom'     ,
    'GE'  , 'Georgia'            ,
    'GR'  , 'Greece'             ,
    'HR'  , 'Croatia'            ,
    'HU'  , 'Hungary'            ,
    'IE'  , 'Ireland'            ,
    'IS'  , 'Iceland'            ,
    'IL'  , 'Israel'             ,
    'IT'  , 'Italy'              ,
    'LT'  , 'Lithuania'          ,
    'LU'  , 'Luxembourg'         ,
    'LV'  , 'Latvia'             ,
    'ME'  , 'Montenegro'         ,
    'MK'  , 'North Macedonia'    ,
    'NL'  , 'Netherlands'        ,
    'NO'  , 'Norway'             ,
    'PL'  , 'Poland'             ,
    'PT'  , 'Portugal'           ,
    'RO'  , 'Romania'            ,
    'RS'  , 'Serbia'             ,
    'RU'  , 'Russian Federation' ,
    'SE'  , 'Sweden'             ,
    'SI'  , 'Slovenia'           ,
    'SK'  , 'Slovakia'           ,
    'TR'  , 'Turkey'             ,
    'UA'  , 'Ukraine'            ,
    'XK'  , 'Kosovo'
  )

ess_11_countries <-
  ess_in |>
  distinct(cntry) |>
  mutate(
    Country = recode_values(
      cntry,
      from = country_coding$Code,
      to = country_coding$Name
    )
  )

country_shape_col <-
  World |>
  filter(continent %in% c("Europe", "Asia")) |>
  full_join(ess_11_countries, by = c("name" = "Country")) |>
  st_crop(st_bbox(c(xmin = -27, ymin = 11, xmax = 60, ymax = 70))) |>
  mutate(ESS11 = !is.na(cntry))

legend_df <- country_shape_col |>
  as_tibble() |>
  filter(ESS11) |>
  arrange(cntry)

tm_shape(country_shape_col) +
  tm_polygons(fill = "lightgrey", col = "darkgrey") +
  tm_shape(country_shape_col |> filter(ESS11)) +
  tm_polygons(fill = "MAP_COLORS") +
  tm_text("cntry")
```

## Accessing ESS data

1. Create an account at the [ESS Data Portal](https://ess.sikt.no/en/register?redirectUrl=/en/) and then get your [user ID](https://ess.sikt.no/en/api)

2. Add your user ID to your R environ file by opening the file and editing as follows:

```r
usethis::edit_r_environ()
```

Edit as follows (example):

```r
ESS_USERID="12345678-1234-1234-1234-12345678900"
```

3. Restart your R session



## Load the ESS data into R

```{r}
#| label: rep-load-ess
#| ref.label: "load-ess"
#| echo: true
#| eval: false
```

## Why weight? Why use the design?

- Differential selection probabilities and nonresponse are accounted for in analytic weights
- Weights are post-stratified by gender, age, education, and geographic region
- Not using weights leads to bias
- Not using design likely leads to incorrect precision, wrong degrees of freedom and not accounting for clustering

## Estimates comparison: weights and design (table)

```{r}
#| label: why-weight
#| echo: false
ess_net_IT <-
  ess_in |>
  filter(cntry == "IT") |>
  select(
    psu,
    cntry,
    stratum,
    anweight,
    pspwght,
    netusoft
  ) |>
  mutate(
    InternetUse = recode_values(
      netusoft,
      1 ~ 'Never',
      2 ~ 'Only occasionally',
      3 ~ 'A few times a week',
      4 ~ 'Most days',
      5 ~ 'Every day'
    ) |>
      fct_reorder(netusoft),
    wt1 = 1
  )

ess_des <- ess_net_IT |>
  as_survey_design(ids = psu, strata = stratum, weights = anweight, nest = TRUE)
ess_des_1 <- ess_net_IT |>
  as_survey_design(ids = psu, strata = stratum, weights = wt1, nest = TRUE)
ess_des_wt <- ess_net_IT |>
  as_survey_design(weights = anweight)
ess_des_srs <- ess_net_IT |>
  as_survey_design(weights = wt1)

calc_ests <- function(des, suff) {
  des |>
    filter(!is.na(InternetUse)) |>
    group_by(InternetUse) |>
    summarize(
      p = survey_prop(proportion = TRUE, prop_method = "logit", vartype = "ci")
    ) |>
    ungroup() |>
    rename(p_est = p) |>
    mutate(p_width = p_upp - p_low) |>
    rename_with(~ str_c(.x, "_", suff), .cols = starts_with("p"))
}

allests <-
  list(a = ess_des, c = ess_des_wt, b = ess_des_1, d = ess_des_srs) |>
  imap(calc_ests) |>
  reduce(full_join, by = "InternetUse")

title_it <- "Estimates of proportion of internet usage in Italy, ESS round 11, under different sample design specifications"

allests |>
  select(-contains("low"), -contains("upp")) |>
  gt() |>
  tab_spanner(
    "Weights, strata, clusters",
    columns = ends_with("a"),
    id = "full"
  ) |>
  tab_spanner(
    "Strata and clusters only",
    columns = ends_with("b"),
    id = "des_only"
  ) |>
  tab_spanner(
    "Weights only",
    columns = ends_with("c"),
    id = "wt_only"
  ) |>
  tab_spanner(
    "No design",
    columns = ends_with("d"),
    id = "no_des"
  ) |>
  cols_label(
    starts_with("p_est") ~ "Proportion",
    starts_with("p_low") ~ "Lower CI",
    starts_with("p_upp") ~ "Upper CI",
    starts_with("p_width") ~ "CI Width",
    InternetUse ~ "netusoft (Internet usage)"
  ) |>
  fmt_percent(decimals = 1) |>
  tab_footnote(
    "95% confidence interval using a Wald-type interval on the log-odds scale.",
    locations = cells_column_labels(c(
      contains("upp"),
      contains("low"),
      contains("width")
    ))
  ) |>
  tab_footnote(
    glue::glue(
      "Degrees of freedom is the number of clusters minus number of stratum which is {survey::degf(ess_des) |> formatC(big.mark=',')}."
    ),
    locations = cells_column_spanners(spanners = c("full", "des_only"))
  ) |>
  tab_footnote(
    glue::glue(
      "Degrees of freedom is the number of respondents minus 1 which is {survey::degf(ess_des_wt) |> formatC(big.mark=',')}."
    ),
    locations = cells_column_spanners(spanners = c("wt_only", "no_des"))
  ) |>
  tab_header(
    title = title_it
  ) |>
  tab_source_note(
    "European Social Survey European Research Infrastructure (ESS ERIC). (2026). ESS11 integrated file, edition 4.2 [Data set]. Sikt - Norwegian Agency for Shared Services in Education and Research. https://doi.org/10.21338/ess11e04_2"
  )
```

## Estimates comparison: weights and design (plot)


```{r}
#| label: weight-compare-plot
#| echo: false
allests |>
  pivot_longer(
    cols = starts_with("p"),
    names_pattern = "p_(.*)_(.*)",
    names_to = c("esttype", "method")
  ) |>
  pivot_wider(names_from = esttype, values_from = value) |>
  mutate(
    Design = recode_values(
      method,
      "a" ~ "Weights, strata, clusters",
      "b" ~ "Strata and clusters only",
      "c" ~ "Weights only",
      "d" ~ "No design"
    ) |> factor(levels=c("Weights, strata, clusters", "Weights only", "Strata and clusters only", "No design"))
  ) |>
  ggplot(aes(color = Design)) +
  geom_point(aes(x = Design, y = est)) +
  geom_errorbar(aes(x = Design, ymin = low, ymax = upp)) +
  facet_wrap(~InternetUse, nrow = 1) +
  theme(
    legend.position = "bottom",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
    labs(title=title_it, y="Estiamted proportion with 95% CI") +
    scale_y_continuous(labels = scales::percent)
```

## ESS Variable Viewer

- Explore ESS variables at the [Variable Viewer](https://ess.sikt.no/en/datafile/242aaa39-3bbb-40f5-98bf-bfb1ce53d8ef?tab=1)

# Closing

## References

---
nocite: |
  @*
---

